MATH 245 Homework 1

### 2024-01-29 PDEs/math245_hw1.Rmd
worked on hw1 PDEs

---
title: "MATH 245 Homework 1"
author: "Ruby Krasnow"
date: "2024-01-29"
output:
  pdf_document:
    extra_dependencies: ["derivative"]
---

@@

### Question 2
(a) 2nd-order linear homogenous  
(b) 4th-order linear inhomogenous  
(c) 2nd-order quasi-linear homogeneous  
(d) We can rewrite this as $u^{-5}u_{xx}+u^{-5}u_{yy}+u^{-4}f(x,y)=g(x,y)$ to see that it is a 2nd-order quasi-linear inhomogeneous PDE.

@@

```{r}
C=1
df_4a <- tibble(x = seq(-2*pi, 2*pi, length.out = 1000),
                y=tan(x+C)) %>% 
  mutate(y = ifelse(abs(y) < 20, y, NA))

@@

```{r}

@@

```{r, warning = FALSE}
curves_4a <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(C*exp(-seq))) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4a)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  ylim(NA, 100)+
  xlim(-5, NA)+
  labs(x="t", y="x", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4b")
```

@@

We check by differentiating:
$$u_t=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)xe^t$$
$$u_x=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)e^t$$
And we see that $u_t-xe_t$ indeed equals zero.
Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:
```{r, warning=FALSE}

@@

We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$.

>>>>

---
title: "MATH 245 Homework 1"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-01-29"
output:
  pdf_document:
    extra_dependencies: ["derivative"]
---

@@

### Question 2
(a) 2nd-order linear homogeneous  
(b) 4th-order linear inhomogeneous  
(c) 2nd-order quasi-linear homogeneous  
(d) We can rewrite this as $u_{xx}+u_{yy}+f(x,y)u-g(x,y)u^5=0$, making it clear that this is a 2nd-order semi-linear homogeneous PDE.

@@

```{r, echo=FALSE}
C=1
df_4a <- tibble(x = seq(-2*pi, 2*pi, length.out = 1000),
                y=tan(x+C)) %>% 
  mutate(y = ifelse(abs(y) < 20, y, NA))

@@

```{r, echo=FALSE}

@@

```{r, warning = FALSE, echo=FALSE}
curves_4b <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(C*exp(-seq))) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4b)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(NA, 100), xlim=c(-5, 9))+
  labs(x="t", y="x", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4b")
```

@@

We check by differentiating:
$$u_t=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)xe^t$$
$$u_x=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)e^t$$
And we see that $u_t-xe_t$ indeed equals zero.
Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:

```{r, warning=FALSE, echo=FALSE}

@@

We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$. This can be written as
\begin{equation}
\lim\limits_{t \to \infty}u(t,x)=
    \begin{cases}
        1 & \text{if } x=0\\
        0 & \text{if } x\neq0
    \end{cases}
\end{equation}

#### (c)
$u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2, \, u(0,y)=e^y$

$\frac{dy}{dx}=x$
Solving this ODE yields the equation for the characteristic curves: $y=\frac{x^2}{2}+C$

```{r, warning = FALSE, echo=FALSE}
curves_4c <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list(0.5*(seq)^2+C)) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4c)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(NA, 45), xlim=c(-9.5, 9.5))+
  labs(x="x", y="y", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4c")
```

Now we define our new coordinate system as $\xi=y-\frac{x^2}{2}, \, \eta=x$

Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
$$\pdv{u}{\eta}-x\pdv{u}{\xi}$$

and similarly
$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}=$$
$$\pdv{u}{\xi}$$

$$u_x+xu_y=\pdv{u}{\eta}-x\pdv{u}{\xi}+x\pdv{u}{\xi}=\pdv{u}{\eta}$$

And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
$$\pdv{u}{\eta}=\left(y-\frac{x^2}{2}\right)^2=\xi^2$$
Now we integrate with respect to $\eta$ to find
$$u=\eta\xi^2+F(\xi)$$

At $u(0,y)$, this becomes $u=0y^2+F(y)$, or $u(0,y)=F(y)$.
On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
$$u=\eta\xi^2+e^\xi$$
Substitute back in $x$ and $y$:
$$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$

We can check by differentiating:
$$u_x=-2x^2\left(y-\frac{x^2}{2}\right)+\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}(-x)$$
$$u_y=2x\left(y-\frac{x^2}{2}\right)+\exp{\left(y-\frac{x^2}{2}\right)}$$

Which means that $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, as we wanted.

#### (d)
$2xu_x+(x+1)u_y=y$, for $x>0, u=2y$ on $x=1$

$$\frac{dx}{dy}=\frac{x+1}{2x}=\frac{1}{2}+\frac{1}{2x}$$
Solving this ODE yields the characteristic curves:
$$y=\frac{x}{2}+\frac{\ln(x)}{2}+C$$
$$C=2y-x-\ln(x)$$

```{r, warning = FALSE, echo=FALSE}
curves_4d <- tibble(C=c(1:5), seq=list(x)) %>% 
  rowwise() %>% 
  mutate(y=list((seq/2)+(log(seq)/2)+C)) %>% 
  unnest(c(seq,y))

ggplot(data=curves_4d)+
  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
  coord_cartesian(ylim=c(-1, 12), xlim=c(-1, 9.5))+
  labs(x="x", y="y", color="C")+
  theme_light()+
  ggtitle("Characteristic curves for Problem 4d")
```

Now we define our new coordinate system as $2y-x-\ln(x), \, \eta=x$

Then by the chain rule,
$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$

$$\pdv{u}{\eta}+\pdv{u}{\xi}\left(-1-\frac{1}{x}\right) $$

Similarly, $$u_y=\pdv{u}{y}=2\pdv{u}{\xi}$$

Plugging into our initial PDE,
$$2xu_x+(x+1)u_y=2x\pdv{u}{\eta}+\pdv{u}{\xi}2x\left(-1-\frac{1}{x}\right)+2(x+1)\pdv{u}{\xi}=2x\pdv{u}{\eta}$$

And since we know $2xu_x+(x+1)u_y=y$, this means $y=2x\pdv{u}{\eta}$ or
$$\frac{\eta}{2}+\frac{\ln(\eta)}{2}+\xi=2\eta \pdv{u}{\eta}$$
